Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)

### 2017-02-16 csf-2017/paper.tex

  \node[main node] (6) [above right of=5]{$W_6$};
  \node[main node] (label6) [above of=6,yshift=-1cm] {\footnotesize{\textit{\texttt{f3} returns}}};
  \path(1) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (2)
       (2) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\pub$} (3)
       (3) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (4)
       (4) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\pub$} (5)
       (5) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (6);
\end{tikzpicture}
\caption{Illustration of the worlds in the proof of
  Lemma~\ref{lem:correctness-g1}. In the proof, the above row of worlds are
  constructed by us, while the below worlds are given. $W_6$ is constructed such
  that $W_6 \futurestr W_5$ \emph{ and } $W_6 \futurewk W_1$.}
  \label{fig:worlds-in-pf-g1}
\end{figure}
\begin{comment}
\begin{itemize}
\item Ticket dispenser
\item The awkward example and variants
\item A sandboxing example?

@@

\emph{Public/private future worlds} A novel aspect of our logical relation is
how we model the special nature of local capabilities using public/private
future worlds. The main insight is that the temporary, revokable nature of these
capabilities can be seen as a generalization of the syntactically-enforced
unstorable status of evaluation contexts in lambda calculi without control
effects. To reason about code that relies on this (particularly, the original
awkward example), \citet{Dreyer:jfp12} (DNB) formally capture the special status
of evaluation contexts using Kripke worlds with public and private future world
relations. Essentially, they allow relatedness of evaluation contexts to be
monotone with respect to a weaker future world relation (public) than
relatedness of values. This formalizes the idea that it is safe to make
temporary internal state modifications (private world transitions, that
invalidate the continuation, but not other values) while a term is performing
internal steps, as long as it returns to a stable state (i.e. transitions to a
public future world of the original) before returning. We generalize this idea
to reason about local capabilities: validity of local capabilities is allowed to
be monotone with respect to a weaker future-world relation than other values,
which we can exploit to distinguish between state changes that are always valid
(public future worlds) and changes that are only valid if we clear local
capabilities (private future worlds). Our future world relations are similar to
\citet{Dreyer:jfp12}'s (for example, our proof of the awkward example uses the
exact same state transition system), but they turn up in an entirely different
place in the logical relation: rather than using public future worlds for the
special syntactic category of evaluation contexts, they are used in the value
relation depending on the locality of the capability at hand. Additionally, our
worlds are more complex because, to allow local memory capabilities and
write-local capabilities, they can contain (revokable) temporary regions that
are only monotonous w.r.t. public future worlds, while DNB's worlds are entirely
permanent.

>>>>

  \node[main node] (6) [above right of=5]{$W_6$};
  \node[main node] (label6) [above of=6,yshift=-1cm] {\footnotesize{\textit{\texttt{f3} returns}}};
  \path(1) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (2)
       (2) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\pub$} (3)
       (3) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (4)
       (4) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\pub$} (5)
       (5) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (6);
\end{tikzpicture}
\caption{Illustration of the worlds in the proof of
  Lemma~\ref{lem:correctness-g1}. In the proof, the top row of worlds are
  constructed by us, while the bottom row of worlds are given. $W_6$ is constructed such
  that $W_6 \futurestr W_5$ \emph{ and } $W_6 \futurewk W_1$.}
  \label{fig:worlds-in-pf-g1}
\end{figure}
\begin{comment}
\begin{itemize}
\item Ticket dispenser
\item The awkward example and variants
\item A sandboxing example?

@@

\emph{Public/private future worlds} A novel aspect of our logical relation is
how we model the special nature of local capabilities using public/private
future worlds. The main insight is that the temporary, revokable nature of these
capabilities can be seen as a generalization of the syntactically-enforced
unstorable status of evaluation contexts in lambda calculi without control
effects (of which well-bracketed control flow is a consequence). To reason about
code that relies on this (particularly, the original awkward example),
\citet{Dreyer:jfp12} (DNB) formally capture the special status of evaluation
contexts using Kripke worlds with public and private future world relations.
Essentially, they allow relatedness of evaluation contexts to be monotone with
respect to a weaker future world relation (public) than relatedness of values,
formalizing the idea that it is safe to make temporary internal state
modifications (private world transitions, which invalidate the continuation,
but not other values) while an expression is performing internal steps, as long
as the code returns to a stable state (i.e. transitions to a public future world
of the original) before returning (i.e. invoking the continuation).\lau{I feel there is an awful lot of parentheses here and that some of them might not be strictly necessary (like explaining temporary internal state modifications and returning = invoking the continuation). Dominique, could you decide whether they are essential to stay true to the contents of the cited paper.}
We
generalize this idea to reason about local capabilities: validity of local
capabilities is allowed to be monotone with respect to a weaker future-world
relation than other values, which we can exploit to distinguish between state
changes that are always safe (public future worlds) and changes that are only
valid if we clear all local capabilities (private future worlds). Our future
world relations are similar to \citet{Dreyer:jfp12}'s (for example, our proof of
the awkward example uses exactly the same state transition system),
but they turn up in an entirely different place in the logical relation: rather
than using public future worlds for the special syntactic category of evaluation
contexts, they are used in the value relation depending on the
locality of the capability at hand. Additionally, our worlds are a bit more
complex because, to allow local memory capabilities and write-local
capabilities, they can contain (revokable) temporary regions that are only
monotonous w.r.t. public future worlds, while DNB's worlds are entirely
permanent.
